---
title: "User based Book Recommender"
subtitle: "Exploratory Analysis & Collaborative Filtering & Shiny App"
output:
html_document:
theme: cosmo
---

Have you ever wondered which book to read next? I often have and to me, book recommendations are a fascinating issue.
This external dataset allows us to take a deeper look at data-driven book recommendations.
THIS IS JUST SELECTION AND IMPUTATION

Let's go.

## Part I: Exploratory Analysis 

### Read in the data
We start by loading some libraries and reading in the two data files.
```{r message=FALSE, warning=FALSE, results='hide'}
library(recommenderlab)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(DT)
library(knitr)
library(grid)
library(gridExtra)
library(corrplot)
library(qgraph)
library(methods)
library(Matrix)

books <- fread('../books.csv')
ratings <- fread('../ratings.csv')
#book_tags <- fread('../book_tags.csv')
#tags <- fread('../tags.csv')
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```



So let's first remove the duplicate ratings. 
```{r}
ratings[, N := .N, .(user_id, book_id)]

## corresponding dplyr code
# ratings %>% group_by(user_id, book_id) %>% mutate(n=n())
cat('Number of duplicate ratings: ', nrow(ratings[N > 1]))
ratings <- ratings[N == 1]

```

And then let's remove users who rated fewer than 5 books. 
```{r}
ratings[, N := .N, .(user_id)]
ratings[, B := .N, .(book_id)]
## corresponding dplyr code
# ratings %>% group_by(user_id) %>% mutate(n = n())
cat('Number of users who rated fewer than 5 books: ', uniqueN(ratings[N <= 5, user_id]))
ratings <- ratings[N > 5]
cat('Number of book who rated fewer than 5 books: ', uniqueN(ratings[B <= 5, book_id]))
ratings <- ratings[B > 5]
cat('Number of users who rated fewer than 5 books g: ', uniqueN(ratings[N <= 5, user_id]))
ratings <- ratings[N > 5]
cat('Number of book who rated fewer than 5 books g: ', uniqueN(ratings[B <= 5, book_id]))
ratings <- ratings[B > 5]
glimpse(ratings)
length(unique(ratings$book_id))
length(unique(ratings$user_id))
```

Setup to remove users have nearly nothing in common with target user. 
```{r}
ratings[, Bmean := mean(rating), .(book_id)]
ratings[, MisMean := signif(rating-Bmean,digits = 3),]
ratings[, shadow := 1,] # then use shadow numbers instead of rating for matix

glimpse(ratings)
ot.use.touch <- ratings[,
        .(sum(B)),
        by = .(user_id)]
bk.ct.b.use <- ratings[,
        .(.N),
        by = .(user_id)]
glimpse(ot.use.touch)
glimpse(bk.ct.b.use)
summary(ot.use.touch)
#ratings[, M := .(mean()), .(book_id)]
#book.mean <- ratings[,
 #       .(mean()),
#        by = .(book_id)]
```

```{r}
# Generate the plot.

p01 <- ot.use.touch %>%
  #dplyr::select(V1) %>%
  ggplot2::ggplot(ggplot2::aes(x=V1)) +
  ggplot2::geom_density(lty=3) +
  ggplot2::xlab("V1\n\nRattle 2018-Mar-15 20:49:55 John") +
  ggplot2::ggtitle("Distribution of V1") +
  ggplot2::labs(y="Density")

# Display the plots.

gridExtra::grid.arrange(p01)
```
```{r}
# Generate the plot.

p01 <- bk.ct.b.use %>%
  #dplyr::select(V1) %>%
  ggplot2::ggplot(ggplot2::aes(x=N)) +
  ggplot2::geom_density(lty=3) +
  ggplot2::xlab("N\n\nRattle 2018-Mar-15 20:49:55 John") +
  ggplot2::ggtitle("Distribution of N") +
  ggplot2::labs(y="Density")

# Display the plots.

gridExtra::grid.arrange(p01)
```


pick the target user
```{r}
set.seed(1)
users <- unique(ratings$user_id)
max(users==17329)
ot.use.touch[ot.use.touch$user_id==17329]
identical(ot.use.touch$user_id,users)
summary(sele.usr<-ot.use.touch$V1>750000)
head(sele.usr)
ot.use.touch[3]
bk.ct.b.use[3]
```
Now that we have selected a user lets remove all books not rated by user.
```{r}
userated<-ratings[user_id==4]
glimpse(userated)
thisway<-(ratings$book_id %in% userated$book_id)
summary(thisway)
useable.rats<-ratings[thisway]
glimpse(useable.rats)
```


And then let's pick number for number of users to remove.
```{r}
useable.rats[, N := .N, .(user_id)]
useable.rats[, B := .N, .(book_id)]
summary(useable.rats$N)
# Generate the plot.
#with(useable.rats, qqPlot(N, dist="norm", id.method="y", id.n=2, 
#  labels=rownames(useable.rats)))

p01 <- useable.rats %>%
  #dplyr::select(V1) %>%
  ggplot2::ggplot(ggplot2::aes(x=N)) +
  ggplot2::geom_density(lty=3) +
  ggplot2::xlab("N\n\nRattle 2018-Mar-15 20:49:55 John") +
  ggplot2::ggtitle("Distribution of Books rated count by user") +
  ggplot2::labs(y="Density")

# Display the plots.

gridExtra::grid.arrange(p01)

# Generate just the data for an Ecdf plot of the variable 'N'.

ds <- rbind(data.frame(dat=useable.rats[,N], grp="All"))

# The 'Hmisc' package provides the 'Ecdf' function.

library(Hmisc, quietly=TRUE)

# Plot the data.

Ecdf(ds[ds$grp=="All",1], col="#E495A5", xlab="N", lwd=2, ylab=expression(Proportion <= x), subtitles=FALSE)


```
Pick the specific number.
```{r}
## corresponding dplyr code
# useable.rats %>% group_by(user_id) %>% mutate(n = n())
cat('Number of users who rated fewer than 20 books: ', uniqueN(useable.rats[N <= 20, user_id]))
useable.rats <- useable.rats[N > 20]


useable.rats[, N := .N, .(user_id)]
useable.rats[, B := .N, .(book_id)]
cat('Number of book who rated fewer than 5 books g: ', uniqueN(ratings[B <= 5, book_id]))
ratings <- ratings[B > 5]
glimpse(useable.rats)
ot.use.touch2 <- useable.rats[,
        .(sum(B)),
        by = .(user_id)]
bk.ct.b.use2 <- useable.rats[,
        .(.N),
        by = .(user_id)]

summary(bk.ct.b.use2)
summary(ot.use.touch2)

```
check for your own ratings and best friends!
```{r}
uquserat<-unique(useable.rats$user_id)
l.uquserat<-length(uquserat)
#userated$rating ; userated$MisMean
correlsp<-vector(mode = 'double',length = max(uquserat))
for(i in uquserat){
  forcor<-merge(userated[,c('book_id','Bmean','MisMean','rating')],
                useable.rats[user_id==i,c('book_id','MisMean','rating')],
                by='book_id',all=T)
  misrates<-is.na(forcor$rating.y)
  forcor$MisMean.y[misrates]<-0
  forcor$rating.y[misrates]<-forcor$Bmean[misrates]
  correlsp[i]=min(cor.test( forcor$MisMean.x, forcor$MisMean.y,conf.level = .95)$conf.int)
}
summary(correlsp)
```
see how many friends you have? and what they recommend?
```{r}
cororder<-order(correlsp, decreasing = T)
correlsp[cororder[200:300]]
#useable.rats[user_id %in% cororder[1:20]]
recc<-ratings[user_id %in% cororder[2:230],.(mean(MisMean)),.(book_id)]
reccorder<-order(recc$V1, decreasing = T)
recc[reccorder[1200:1220]]
```
```{r}
compare<-merge(useable.rats[user_id==4],recc,by='book_id',all=T)
#compare[user_id==4,c('rating','V1')][1:134]
cor.test(compare$MisMean,compare$V1)
RMSE(compare$MisMean,compare$V1)
vecmean<-vector()
for(i in 1:134){ vecmean[i]<-mean(compare$MisMean, na.rm=T)}
cor.test(userated$MisMean,vecmean)
RMSE(userated$MisMean,vecmean)
cor.test(compare$MisMean,compare$V1)
cor.test(ratings[user_id==4]$rating,ratings[user_id==4]$Bmean)
cor.test(ratings[user_id==4]$rating,ratings[user_id==4]$MisMean)
cor.test(ratings[user_id==4]$Bmean,ratings[user_id==4]$MisMean)
glimpse(ratings)
str(ratings[user_id==4,'rating'])
```
```{r}
library(Matrix)
ratings
res <- prcomp(Xt, center = TRUE, scale = FALSE)
plot(cumsum(res$sdev^2/sum(res$sdev^2))) #cumulative explained variance
```
Because this matrix has many columns, I would like to reduce its dimensionality to something more manageable. I can use the excellent irlba package to perform SVD and return the first n principal components (5 shown here; I'll probably use 100 or 500 on my actual dataset):
```{r}
library(irlba)
 pc <- M %*% irlba(M, nv=5, nu=0)$v 
 str(pc)
```





```{r}
# Generate the plot.

p01 <- ot.use.touch2 %>%
  #dplyr::select(V1) %>%
  ggplot2::ggplot(ggplot2::aes(x=V1)) +
  ggplot2::geom_density(lty=3) +
  ggplot2::xlab("V1\n\nRattle 2018-Mar-15 20:49:55 John") +
  ggplot2::ggtitle("Distribution of Touch by user") +
  ggplot2::labs(y="Density")

# Display the plots.

gridExtra::grid.arrange(p01)
```
```{r}

```
melt spread and butter tuple dataset into matrix
```{r}
ratings<-useable.rats
dimension_names <- list(user_id = sort(unique(ratings$user_id)), book_id = sort(unique(ratings$book_id)))
if(F){
  ratingmat <- spread(select(ratings, book_id, user_id, MisMean), book_id, MisMean) %>%     select(-user_id)
}else{
  ratingmat <- spread(select(ratings, book_id, user_id, shadow), book_id, shadow) %>% select(-user_id)
}
ratingmat <- base::as.matrix(ratingmat)
dimnames(ratingmat) <- dimension_names
ratingmat[1:5, 1:5]
dim(ratingmat)
```
Print
```{r}
lrm<-length(ratingmat[,1])
if(T){ratingmat[is.na(ratingmat)]<-0}
str(ratingmat)
if(F){ratingmat<-data.table(useable.rats[user_id==4, MisMean], 
                        ratingmat)
write.table(t(ratingmat[c(1,3,2,4:lrm),]),file = "4thbookcentqrtshad.csv",quote = F,sep = ",",row.names = F ,col.names = F)
}else{
  ratingmat<-data.table(useable.rats[user_id==4, rating], 
                        ratingmat)
write.table(t(ratingmat[c(1,3,2,4:lrm),]),file = "4thbookqrtshad.csv",quote = F,sep = ",",row.names = F ,col.names = F)
}
write.table(t(ratingmat[c(2,1,3:lrm),]),file = "4thbookcentqrtshad.csv",quote = F,sep = ",",row.names = F ,col.names = F)
```

```{r}
rm(books, ratings,useable.rats,bk.ct.b.use,bk.ct.b.use2,ot.use.touch,ot.use.touch2)
gc()
```
